Setup Honeypot: Team Contagiarium

### 2011-11-01 Setup_honeypot.tex
Auditing/monitoring bit

@@

To be able to still log on to the honeypot, we configured SSH to use port 9022.

@@

\subsection{Snort}
Snort is an intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
Snort was designed to be used as a network IDS, however we will be running it on our host machine, effectively making it a host-based IDS.

Unfortunately the VM network went down before we were able to complete our Snort configuration.
We intend to set up Snort using a predefined rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
The rules that we will be using are based on a honeypot rule-set. % TODO reference
Included are rules on detecting backdoors, viruses, DDOS, spam and other abuse.
These rules will be tweaked if the results are unsatisfactory.
If a rule is triggered, snort will log an alert.

We will configure Snort to send alerts to syslog, and to log its packet captures to some form of read-only filesystem mounted from our monitoring machine (yet to be decided).
We will install a script like \verb|auto-oink| to automatically retrieve the latest community Snort rules.

>>>>

\subsection{Monitoring}
After collecting our syslog information, we want to somehow be alerted of interesting information automatically.
Some seemingly appropriate projects are available under an open source license, and we have not yet decided which one we want to use.
A customizable solution is absolutely necessary owing to the wide range of incoming information and the modifications made to a few of the installed software packages specifically for logging to syslog.
Next to any off-the-shelf solution, some more custom scripting seems inevitable here.

@@

To be able to still log on to the honeypot, we configured our real SSH service to use port 9022.

We have patched Kippo to log messages to syslog.

@@

\subsection{Snort}
Snort is an open source intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
Snort was designed to be used as a network IDS, however we will be running it on our host machine, effectively making it a host-based IDS.

Unfortunately the VM network went down before we were able to complete our Snort configuration.
We intend to set up Snort using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
Included are rules on detecting exploit attempts on various common services like HTTP, FTP, SMTP, and many more.

Considering we are in a VM we might be limited in RAM, which Snort seems to require a lot of.
The configured rules will have to be minimal to avoid exhausting resources.

We will configure Snort to send alerts to syslog, and to log its packet captures to a read-only \verb|sshfs| mounted from our monitoring machine.
We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.

\subsection{Auditing}
For the HI phase, we would like to have a way to audit commandline interaction. 
Various options exist, such as the user-space \verb|auditd| daemon, but to make auditing harder to detect this we feel it would be best to look for a kernel-space solution.
An auditing module developed by the Honeynet project called Sebek can provide this, but is unmaintained and outdated.
The grsec kernel patches also provide auditing from kernel space, but we were advised against replacing the stock kernel.

We are still investigating the best way to achieve useful auditing.
A self-written kernel module that hooks system calls, albeit limited, might suffice to get some much needed auditing information from the honeypot.
